Supplement to Automatic Matching of Bullet Lands

### 2016-01-07 imaging-paper.Rnw
couple more words on the extended study

set.seed(20151202)
names(bstats)[9:10] <- c("num.matches", "num.nonmatches")
includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))

@@

Another benefit of the digitized version of the images is that we can not only apply a single decision tree, but several hundreds of them to combine in a random forest~\citep{breiman:2001, randomForest}.  For each of the trees in a random forest, only two thirds of the observations are used for fitting, while  the remaining third  are used to evaluate the tree's predictive power and accuracy, or its reverse, the error rate. Because errors are determined from the one third of held-back observations, this error rate is called the out-of bag (OOB) error. 
Figure~\ref{fig:oob} shows the cumulative out-of-bag error (OOB) rate for 300 trees. 
%
\begin{figure}[hbtp]
  \centering
<<oob, echo=FALSE, fig.width=6, fig.height=3, out.width='.6\\textwidth'>>=
qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
@
\caption{\label{fig:oob}Cumulative out-of-bag error rate of a random forest fit to predict land-to-land matches from image features.}
\end{figure}
%
After about 100 trees, the error rate of land-to-land comparisons stabilizes at 0.0039. This is a weighted average between false positive error rate of 0.0002 and an error rate of false negatives of 0.2267. This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:tree-forest}).
Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches. In the case of the Hamby study, even for the single tree the overall error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched. % (see Figure~\ref{fig:treeresult}). 
%' \begin{figure}[hbtp]
%'   \centering
%' <<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
%' bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
%' bullets <- bstats %>% filter(match, !flagged) %>% group_by(bullet) %>% summarize(
%'   n = n(),
%'   pred = sum(pred > 0.5),
%'   forest = sum(forest > 0.5)
%' )
%' library(reshape2)
%' bm <- melt(data=bullets, id.var=c("bullet", "n"))
%' 
%' bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
%' qplot(data=bm, 100*value/n, reorder(bullet, value/n), size=I(2.5), 
%'       shape=prediction, colour=prediction, alpha=I(0.6)) + 
%'   theme_bw()  + ylab("") + 
%' #  scale_shape_manual(values=c(1,2)) +
%'   scale_x_continuous("Percent of correctly predicted land-to-land matches", 
%'                      breaks = c(0,25,50,75,100), limits=c(0,100)) +
%'   theme(legend.position = "bottom")
%' @
%' \caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12, except for Q and B, where the maximum is 11). Both results for the tree and the random forest are shown. In the tree, at least \hh{40\% of all possible} lands of a bullet are  matched correctly, while the random forest achieves perfect results. Unknown bullet B is the hardest for the tree to match.}
%' \end{figure}
%' %
This makes the errors in the automated approach smaller than the human error in the Hamby study. Out of the 507 participants, who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, resulting in a false negative rate of 0.0011.

>>>>

set.seed(20160105)
names(bstats)[9:10] <- c("num.matches", "num.nonmatches")
includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))

@@

Another benefit of the digitized version of the images is that we can not only apply a single decision tree, but several hundreds of them to combine in a random forest~\citep{breiman:2001, randomForest}.  For each of the trees in a random forest, only two thirds of the observations are used for fitting, while  the remaining third  are used to evaluate the tree's predictive power and accuracy, or its reverse, the error rate. Because errors are determined from the one third of held-back observations, this error rate is called the out-of bag (OOB) error. 
Figure~\ref{fig:oob} shows the cumulative out-of-bag error (OOB) rate for 300 trees. 
%
\begin{figure}[hbtp]
  \centering
<<oob, echo=FALSE, fig.width=6, fig.height=3, out.width='.6\\textwidth'>>=
qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
@
\caption{\label{fig:oob}Cumulative out-of-bag error rate of a random forest fit to predict land-to-land matches from image features.}
\end{figure}
%
After about 100 trees, the error rate of land-to-land comparisons stabilizes at 0.0039. This is a weighted average between false positive error rate of 0.0001 and an error rate of false negatives of 0.2267. This out-of-bag error rate is over-estimating the actual error in the Hamby study: here, the final random forest based on 300 trees is able to correctly predict all known  matches and non-matches (see Figure~\ref{fig:tree-forest}).
Note that this error rate is based on land-to-land comparisons and is much lower for bullet-to-bullet matches. In the case of the Hamby study, even for the single tree the overall error rate is zero, if we require for a bullet match that at least two of a bullet's lands are matched. % (see Figure~\ref{fig:treeresult}). 
%' \begin{figure}[hbtp]
%'   \centering
%' <<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
%' bstats$bullet <- gsub("-[0-9]$", "", bstats$b2)
%' bullets <- bstats %>% filter(match, !flagged) %>% group_by(bullet) %>% summarize(
%'   n = n(),
%'   pred = sum(pred > 0.5),
%'   forest = sum(forest > 0.5)
%' )
%' library(reshape2)
%' bm <- melt(data=bullets, id.var=c("bullet", "n"))
%' 
%' bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
%' qplot(data=bm, 100*value/n, reorder(bullet, value/n), size=I(2.5), 
%'       shape=prediction, colour=prediction, alpha=I(0.6)) + 
%'   theme_bw()  + ylab("") + 
%' #  scale_shape_manual(values=c(1,2)) +
%'   scale_x_continuous("Percent of correctly predicted land-to-land matches", 
%'                      breaks = c(0,25,50,75,100), limits=c(0,100)) +
%'   theme(legend.position = "bottom")
%' @
%' \caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12, except for Q and B, where the maximum is 11). Both results for the tree and the random forest are shown. In the tree, at least \hh{40\% of all possible} lands of a bullet are  matched correctly, while the random forest achieves perfect results. Unknown bullet B is the hardest for the tree to match.}
%' \end{figure}
%' %
This makes the errors in the automated approach smaller than the human error in the Hamby study. Out of the 507 participants, who returned results, eight (out of $15 \times 507 = 7,605$) bullets were not matched conclusively, resulting in a false negative rate of 0.0011.

### 2016-01-07 imaging-supplement.Rnw
couple more words on the extended study

\section{Cylindrical Fit}\label{supp:cylindrical}
Figure~\ref{fig:fixedX} shows the side profile of a crosscut through the surface measurements of bullet 1-5 at a fixed level. The smooth line on top is a circle, with estimated radius and center. The details of this fit are given below. We then remove the fit to investigate residuals.
%
\begin{figure}[hbtp]
  \centering
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\textwidth'>>=

@@

Assume that $n$ data points are given in the form of data tuples $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ that are (approximately) located on a circle. We want to estimate the location of the center and radius of the best fitting circle using a least squares approach.

@@

\section{Signature intensities}\label{supp:bulletbottom}
<<setup-signatures, echo=FALSE>>=
library(x3pr)
library(x3prplus)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(zoo)

@@

\newpage
\section{Complete evaluation of the Hamby study}\label{supp:extended}
\hh{One way to expand the use of the James Hamby study is to not only match all of the unknown bullet lands against the known bullet lands, but to compare every land against every other land. This effectively doubles the number of comparisons from 10,384 pairwise comparisons of usable bullet lands to 21,115 $\left[= (118+88)\cdot 205/2\right]$ comparisons by adding another 10,731 bullet land comparisons made up of known-to-known and unknown-to-unkown comparisons. }

\hh{When we predict the new 10,731 comparisons using the random forest based on the previously considered 10,384 known-unknown comparisons, we encounter 18 false negatives and 9 false positives, corresponding to an actual false error rate of 0.19 and a false positive rate of 0.00085, which is close to the estimated error rates of 0.2267 and 0.0002. }

\hh{However, if we use all of the available comparisons to fit another random forest of 300 trees, the de-facto error rates for false positives and false negatives are again at 0. The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate.}
<<allhamby, echo=FALSE, message=FALSE, warning=FALSE>>=
knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
knowns <- dir(path=knowndatadir)
knowns <- gsub(".x3p", "", knowns)

>>>>

\section{Cylindrical Fit}\label{supp:cylindrical}
Figure~\ref{fig:fixedX} shows the side profile of a crosscut through the surface measurements of bullet 1-5 at a fixed level. The smooth line on top is a circle, with estimated radius and center. The details of this fit are given below:
%
\begin{figure}[hbtp]
  \centering
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\textwidth'>>=

@@

Assume that $n$ data points are given in the form of data tuples $(x_1, y_1)$, $(x_2, y_2)$, $...$, $(x_n, y_n)$ that are (approximately) located on a circle. We want to estimate the location of the center and radius of the best fitting circle using a least squares approach.

@@

\newpage
\section{Signature intensities}\label{supp:bulletbottom}
<<setup-signatures, echo=FALSE>>=
library(x3pr)
library(x3prplus)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(zoo)

@@

%\newpage
\section{Complete evaluation of the Hamby study}\label{supp:extended}
\hh{One way to expand the use of the James Hamby study is to not only match all of the unknown bullet lands against the known bullet lands, but to compare every land against every other land. This effectively doubles the number of comparisons from 10,384 pairwise comparisons of usable bullet lands to 21,115 $\left[= (118+88)\cdot 205/2\right]$ comparisons by adding another 10,731 bullet land comparisons made up of known-to-known and unknown-to-unkown comparisons. }

\hh{When we predict the new 10,731 comparisons using the random forest based on the previously considered 10,384 known-unknown comparisons, we encounter 18 false negatives and 9 false positives, corresponding to an actual false error rate of 0.19 and a false positive rate of 0.00085, which is close to the random forest's estimated OOB error rates of 0.226744 and 0.000098. }

\hh{However, if we use all of the available comparisons to fit another random forest of 300 trees, the de-facto error rates for false positives and false negatives are again at 0. The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate. The false positive rate is therefore virtually unchanged, while we see a slight improvement in the false negative rate for an overall OOB error rate of 0.3\%.}
<<allhamby, echo=FALSE, message=FALSE, warning=FALSE>>=
knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
knowns <- dir(path=knowndatadir)
knowns <- gsub(".x3p", "", knowns)
